---
output:
   github_document:
     pandoc_args: --webtex
---

```{r setup, include=FALSE}

knitr::opts_knit$set(root.dir = "~/RProjects/StripedBassPopMetrics/")

knitr::opts_chunk$set(echo = FALSE)

now <- Sys.Date()

```

```{r load-libraries}

library(sportfish)
# library(package)

```

```{r load-data}

# the data directory for bay study
data_dir <- "data/tagging"

# Tagging <- new.env()
# ReadRDSFiles(fileDir = data_dir, envir = Tagging)

Stripers <- readRDS(file = file.path(data_dir, "StripedBass.rds"))
Effort <- readRDS(file = file.path(data_dir, "Effort.rds"))

# clean up
rm(data_dir)

```

## Background

The California Department of Fish & Wildlife (CDFW) studies adult Striped Bass using mark-recapture (tagging) methods. Since 1969, CDFW has used gill nets and fyke traps to capture and then disc tag many Striped Bass

...
more to come

## Effort

```{r effort}

split_effort <- split(Effort, f = Effort[["RelYear"]])

# to get annual summary (e.g., days fished; traps fished)
effort <- vapply(
  split_effort,
  FUN = EffortSb,
  FUN.VALUE = as.list(numeric(7L)),
  sTime = StartTime, eTime = EndTime, netNum = DriftTrap
)

# next steps for convenience of display
effort <- data.frame(
  Year = as.numeric(dimnames(effort)[[2]]),
  t(effort),
  row.names = NULL
)

effort[] <- lapply(effort, FUN = unlist, use.names = FALSE)

# not really needed (& field is somewhat confusing)
effort$NetCountDup <- NULL

```

Here we display --- for now --- some annual effort for fyke traps. Data exist for other years and for gill net operations but are not readily available in electronic form. In the future, we hope to make available these data. Field headings are explained below.

- Year: calendar year in which sampling was performed
- Total: total hours fyke traps fished
- Avg: daily average hours fyke traps fished
- Var: daily variance hours fyke traps fished
- Days: number of days fished
- Fished: total (seasonal) number of fyke traps fished
- Count: maximum daily number of fyke traps fished

```{r effort-display}

knitr::kable(
  effort,
  format = "markdown",
  digits = 3,
  col.names = c(
    "Year", "Total", "Avg", "Var",
    "Days", "Fished", "Count"
  )
)

```

```{r days-gear}

# un-comment as needed
# head(Stripers, n = 2)
# table(Stripers$CapMethod, useNA = "ifany")

# to get days fished per year by gear type
days_gear <- with(data = Stripers, expr = {
  
  # to produce 0s in years of no sampling
  l <- min(RelYear):as.numeric(format(Sys.Date(), format = "%Y"))
  
  yr <- factor(RelYear, levels = l)
  
  cm <- factor(CapMethod, levels = 1:2, labels = c(`1` = "GN", `2` = "FT"))
  
  s <- split(RelDate, f = list(yr, cm), drop = FALSE)
  
  vapply(s, FUN = function(x) {
    length(unique(x))
  }, FUN.VALUE = numeric(1L))
  
})

# for convenience of display
days_gear <- data.frame(
  do.call(
    what = rbind,
    args = strsplit(names(days_gear), split = "\\.")
  ),
  Days = days_gear,
  row.names = NULL,
  stringsAsFactors = FALSE
)

# did not bother change column names
# X1 = year
# X2 = gear type

# for convenience of plotting
days_gear$X1 <- as.numeric(days_gear[["X1"]])

```

```{r years-no-sampling}

# for convenience in narrative
years_no_sampling <- with(data = days_gear, expr = {
  
  n <- table(X1[Days == 0])
  out <- names(n)[n == 2]
  
  paste0(out, collapse = "; ")
  
})

```

Here we display annual time series (since `r min(Stripers[["RelYear"]])`) of days sampled (fished) by gear type (GN = gill net; FT = fyke trap). We did not conduct mark-recapture operations in the following years: `r years_no_sampling`.

*Note*: We essentially stopped gill netting (GN) in 2009. We deployed gill nets only one day in 2011.

```{r plot-days-gear}

# for stacked plot
mat_layout <- matrix(data = 1:2, nrow = 2, ncol = 1, byrow = TRUE)
nf <- layout(mat = mat_layout)
# layout.show(n = nf)

# for keeping bottom & left spaces to display axes title
# par(oma = c(4, 5, 1, 1), cex.axis = 1.5, family = "sans")
par(oma = c(3, 3, 0.5, 1), cex.axis = 1.05)

with(data = days_gear, expr = {
  
  # to create empty plot
  rngx <- range(X1)
  rngy <- range(Days)
  
  # for plotting both gear types
  vals <- lapply(list("GN", "FT"), FUN = function(g) {
    
    par(mar = c(0.1, 0.5, 1.0, 0.1), mgp = c(1, 0.6, 0))
    
    # to get data specific to gear type
    b <- X2 %in% g
    
    # to plot annual ticks every 5 years
    xtick <- X1 %% 5 == 0
    
    plot(
      x = rngx,
      y = rngy,
      type = "n",
      xaxt = "n",
      las = 1,
      yaxt = "n",
      # xlab = "Year",
      # ylab = sprintf(fmt = "Days sampled (%s)", g),
      xlim = rngx + c(1, -1)
    )
  
    # custom y-axis tick display
    axis(
      side = 2,
      at = axTicks(side = 2)[c(TRUE, FALSE)],
      labels = axTicks(side = 2)[c(TRUE, FALSE)],
      col = "transparent",
      col.ticks = "grey30",
      cex = 1.05,
      las = 1
    )
  
    # for grey background & white gridlines
    par(xaxp = c(rngx, diff(rngx)))
    grid(lwd = 1000, col = "grey90")
    grid(lty = 1, col = "white", lwd = 1)
  
    # plot data as bar
    lines(
      # x = X1[b] - 0.2,
      x = X1[b],
      y = Days[b],
      lend = 1,
      lwd = 6,
      # col = "orange2",
      # col = rgb(0, 0, 0, 0.75),
      col = "grey40",
      type = "h"
    )
    
    # for identifying which plot is which
    mtext(
      text = sprintf(fmt = "Gear type: %s", g),
      side = 3,
      line = 0,
      adj = 0
    )
    
    # output for convenience of custom x-axis
    xtick
  })
  # end lapply()
  
  # resent to one plot
  layout(mat = 1)
  
  # custom x-axis
  axis(
    side = 1,
    at = X1[vals[[1]]],
    labels = X1[vals[[1]]],
    col = "transparent",
    col.ticks = "grey30",
    cex = 1.05
  )
  
  # axis labels
  mtext(text = "Year", side = 1, line = 1.75)
  mtext(text = "Days sampled", side = 2, line = 2)
})

```

## Catch

```{r catch}

# for convenience of splitting & renaming
split_vars <- c("RelYear", "CapMethod")

# to get at count tagged & total catch per split_vars
split_striper <- split(
  Stripers[c("TagNum", "TagAction", "Count")], 
  f = Stripers[split_vars],
  drop = FALSE
)

# total catch & number tagged
catch <- vapply(split_striper, FUN = function(x) {
  
  # because tagging action 1 = tagged & field only active since 2011 not since
  # inception 1969 (prior to 2011 use NA)
  b <- is.na(x[["TagAction"]]) | x[["TagAction"]] %in% 1
  
  c(Tagged = sum(b), Total = sum(x[["Count"]]))
  
}, FUN.VALUE = numeric(2L))

# for convenience of display
catch <- data.frame(
  do.call(
    what = rbind,
    args = strsplit(dimnames(catch)[[2]], split = "\\.")
  ),
  t(catch),
  stringsAsFactors = FALSE,
  row.names = NULL
)

catch[] <- lapply(catch, FUN = as.numeric)

colnames(catch)[1:length(split_vars)] <- split_vars

```

```{r catch-display, eval=FALSE}

# table is quite long to display, opting for graphic instead

catch$CapMethod <- factor(
  catch[["CapMethod"]],
  levels = 1:2,
  labels = c(`1` = "GillNet", `2` = "Fyke")
)

knitr::kable(
  reshape(
    data = catch,
    timevar = "CapMethod",
    idvar = "RelYear",
    direction = "wide",
    sep = " "
  ),
  format = "markdown"
)

```

Here we plot annual time series by gear type (GN = gill net, FT = fyke trap) of total Striped Bass caught and Striped bass tagged (of those caught).

<!-- *Note*: <need to talk about how earlier years electronic data not available for fish not tagged> -->

```{r plot-catch}

# head(catch)

# for stacked plot
mat_layout <- matrix(data = 1:2, nrow = 2, ncol = 1, byrow = TRUE)
nf <- layout(mat = mat_layout)
# layout.show(n = nf)

# for keeping bottom & left spaces to display axes title
# par(oma = c(4, 5, 1, 1), cex.axis = 1.5, family = "sans")
par(oma = c(3, 3, 0.5, 1), cex.axis = 1.05)

with(data = catch, expr = {
  
  # to create empty plot
  rngx <- range(RelYear)
  rngy <- range(Total)
  
  # 
  all_years <- rngx[1]:rngx[2]
  xtick <- all_years %% 5 == 0
  # print(xtick)
  
  # for plotting both gear types
  vals <- lapply(list(1, 2), FUN = function(g) {
    
    par(mar = c(0.1, 0.5, 1.0, 0.1), mgp = c(1, 0.6, 0))
    
    # to get data specific to gear type
    b <- CapMethod %in% g & Total != 0
    
    # to plot annual ticks every 5 years
    # xtick <- all_years %% 5 == 0
    
    plot(
      x = rngx,
      y = rngy,
      type = "n",
      xaxt = "n",
      las = 1,
      yaxt = "n",
      # xlab = "Year",
      # ylab = sprintf(fmt = "Days sampled (%s)", g),
      xlim = rngx + c(1, -1)
    )
    
    # print(axTicks(side = 2))
    # y_format <- AxisFormat(axTicks(side = 2))
  
    # custom y-axis tick display
    axis(
      side = 2,
      at = axTicks(side = 2),
      labels = axTicks(side = 2) / 1000,
      col = "transparent",
      col.ticks = "grey30",
      cex = 1.05,
      las = 1
    )
  
    # for grey background & white gridlines
    par(xaxp = c(rngx, diff(rngx)))
    grid(lwd = 1000, col = "grey90")
    grid(lty = 1, col = "white", lwd = 1)
  
    # plot total data as bar
    lines(
      # x = X1[b] - 0.2,
      x = RelYear[b],
      y = Total[b],
      lend = 1,
      lwd = 10,
      # col = "orange2",
      # col = rgb(0, 0, 0, 0.75),
      col = "grey40",
      type = "h"
    )
    
    # plot tagged data as bar
    lines(
      # x = X1[b] - 0.2,
      x = RelYear[b],
      y = Tagged[b],
      lend = 1,
      lwd = 1,
      # col = "orange2",
      # col = rgb(0, 0, 0, 0.75),
      col = "orange2",
      type = "h"
    )
    
    glkp <- c(`1` = "GN", `2` = "FT")
    
    # for identifying which plot is which
    mtext(
      text = sprintf(fmt = "Gear type: %s", glkp[g]),
      side = 3,
      line = 0,
      adj = 0
    )
    
    # output for convenience of custom x-axis
    # xtick
    sum(b)
  })
  # end lapply()
  
  # resent to one plot
  layout(mat = 1)
  # print(vals)
  
  # custom x-axis
  axis(
    side = 1,
    # at = RelYear[vals[[1]]],
    # labels = RelYear[vals[[1]]],
    at = all_years[xtick],
    labels = all_years[xtick],
    col = "transparent",
    col.ticks = "grey30",
    cex = 1.05
  )
  
  # axis labels
  mtext(text = "Year", side = 1, line = 1.75)
  mtext(text = bquote("Count" ~~ 'x' ~ 10^3), side = 2, line = 2)
  
  # print(axTicks(side = 1))
  
  par(lend = 1)
  
  legend(
    x = 2018,
    y = 15200,
    legend = c("Total", "Tagged"),
    col = c("grey40", "orange2"),
    lwd = c(10, 1),
    border = NA,
    # col = NA,
    bty = "n",
    ncol = 2,
    xpd = TRUE,
    xjust = 1,
    yjust = 0,
    x.intersp = 0.5,
    seg.len = 2
  )
})

```

---

Contact [Jason DuBois](mailto:jason.dubois@wildlife.ca.gov)  
CDFW, SportFish Unit    
`r Sys.Date()`
