---

---

```{r setup0, include=FALSE}

knitr::opts_knit$set(root.dir = "~/RProjects/StripedBassPopMetrics/")

```

<!-- Sampling Frequency -->
<!-- Sampling Period (why chosen) -->
<!-- Sampling Locations -->
<!-- Describe how sampling has changed over time -->
<!-- Plot coordinates?? -->

```{r tag-values}

# lkp_striper$tagvalue
# names(lkp_striper$tagvalue)

tag_values <- with(data = Tagging[["StripedBass"]], expr = {
  
  b <- TagAction %in% 1 | is.na(TagAction)
  
  tv <- factor(TagValue[b], levels = names(lkp_striper[["tagvalue"]]))
  
  table(RelYear[b], tv, useNA = "ifany", dnn = NULL)
  
})

```

```{r rel-cond-annual, eval=FALSE}

# lkp_striper$condition

with(data = Tagging[["StripedBass"]], expr = {
  
  cnd <- lkp_striper[["condition"]]
  
  b <- TagAction %in% 1 | is.na(TagAction)
  
  rc <- factor(RelCond[b], levels = names(cnd), labels = cnd)
  
  table(RelYear[b], rc, useNA = "ifany", dnn = NULL)
  
})

```

```{r effort}

# table(Tagging$Effort$CapMethod, useNA = "ifany")

effort <- Split(
  data = Tagging[["Effort"]],
  # subset = ,
  vars = -RelYear,
  splitVars = RelYear
)

# for desired data type
effort$RelYear <- as.numeric(effort[["RelYear"]])

```

```{r effort-hours}

effort$Hours <- t(vapply(effort[["Data"]], FUN = function(d) {
  
  # for correct number of values in inner vapply & name assignment
  n <- 6L
  
  # so both gear types are summarized
  d$CapMethod <- factor(
    d[["CapMethod"]],
    levels = c(2, 1),
    labels = c("FT", "GN")
  )
  
  # split on gear type (capture method)
  s <- Split(data = d, splitVars = CapMethod)
  
  r <- vapply(s[["Data"]], FUN = function(dd) {
    
    o <- EffortSb(
      data = dd,
      sTime = StartTime,
      eTime = EndTime,
      netNum = DriftTrap
    )
    
    # unlists & removes NetCountDup field
    c(o, recursive = TRUE)[-7]
    
  }, FUN.VALUE = numeric(n))
  
  # creates names for unlisted `r`
  dn <- dimnames(r)
  r <- c(r, recursive = TRUE)
  names(r) <- paste(dn[[1]], rep(dn[[2]], each = n), sep = "")
  r
  
}, FUN.VALUE = numeric(12L)))

```

```{r catch}

# for splitting to ensure each capture method present in every release year
Tagging$StripedBass <- within(data = Tagging[["StripedBass"]], expr = {
  CapMethod2 <- factor(CapMethod, levels = c(2, 1), labels = c("FT", "GN"))
})

# for quick check to confirm above
# with(data = Tagging[["StripedBass"]], expr = {
#   table(CapMethod, CapMethod2, useNA = "ifany")
# })

catch <- Split(
  data = Tagging[["StripedBass"]],
  # subset = ,
  vars = -c(RelYear, CapMethod2),
  splitVars = c(RelYear, CapMethod2)
)

# for desired data type
catch$RelYear <- as.numeric(catch[["RelYear"]])

# for code built using `CapMethod` not `CapMethod2`
colnames(catch) <- sub(pattern = "[0-9]", replacement = "", x = colnames(catch))

```

```{r catch-count}

catch$Count <- t(vapply(catch[["Data"]], FUN = function(d) {
  b <- d[["TagAction"]] %in% 1 | is.na(d[["TagAction"]])
  # sum(d[b, "Count"])
  c(Tagged = sum(d[b, "Count"]), Total = sum(d[["Count"]]))
}, FUN.VALUE = numeric(2L)))

```

```{r catch-days-loc}

catch$DaysLoc <- t(vapply(catch[["Data"]], FUN = function(d) {
  
  d$RelLoc <- factor(d[["RelLoc"]], levels = 1:16)
  s <- Split(data = d, vars = RelDate, splitVars = RelLoc)
  vapply(s[["Data"]], FUN = function(x) {
    length(unique(x[["RelDate"]]))
  }, FUN.VALUE = numeric(1L))
  
}, FUN.VALUE = numeric(16L)))

```

```{r sex-ratio}

sex_ratio_overall <- with(data = Tagging[["StripedBass"]], expr = {
  SexRatio(lkp_striper$sex(RelSex))
})

catch$SexRatio <- t(vapply(catch[["Data"]], FUN = function(d) {
  s <- lkp_striper$sex(d[["RelSex"]], asFactor = TRUE)
  r <- c(SexRatio(s), recursive = TRUE)
  names(r) <- sub(pattern = "\\.", replacement = "", x = names(r))
  r
}, FUN.VALUE = numeric(6L)))

```

```{r rel-cond}

# release condition for disc tagged fish only

catch$RelCond <- t(vapply(catch[["Data"]], FUN = function(d) {
  
  cnd <- lkp_striper[["condition"]]
  
  b <- d[["TagAction"]] %in% 1 | is.na(d[["TagAction"]])
  
  rc <- factor(
    d[b, "RelCond"],
    levels = c(names(cnd), NA),
    labels = c(cnd, "unk"),
    exclude = NULL
  )
  
  table(rc, dnn = NULL)
  
}, FUN.VALUE = numeric(3L)))

```

<!-- catch per date per trap (for CPUE) -->

```{r catch-date-trap}

catch$DateDriftTrap <- lapply(catch[["Data"]], FUN = function(d) {
  
  if (nrow(d) == 0) return(NULL)
  
  d$DriftTrap[is.na(d[["DriftTrap"]])] <- 0
  d$TagAction[is.na(d[["TagAction"]])] <- "1"
  
  dd <- Split(
    data = d,
    vars = c(TagAction, Count),
    splitVars = c(RelDate, DriftTrap)
  )
  
  dd$RelDate <- as.POSIXct(dd[["RelDate"]])
  
  dd$Total <- vapply(dd[["Data"]], FUN = function(ddd) {
    sum(ddd[["Count"]])
  }, FUN.VALUE = numeric(1))

  # dd[c("RelDate", "DriftTrap", "Total")]
  dd$Data <- NULL
  
  dd
})

```

```{r catch-date}

catch$PerDate <- lapply(catch[["Data"]], FUN = function(d) {
  
  if (nrow(d) == 0) return(NULL)
  
  dd <- Split(data = d, vars = Count, splitVars = RelDate)
  
  dd$RelDate <- as.POSIXct(dd[["RelDate"]])
  
  dd$Total <- vapply(dd[["Data"]], FUN = function(ddd) {
    sum(ddd[["Count"]])
  }, FUN.VALUE = numeric(1))
  
  dd$Data <- NULL
  
  dd
})

```

```{r flen-stats, warning=FALSE}

catch$LenStats <- t(vapply(catch[["Data"]], FUN = function(d) {
  
  b <- !d[["RelFLFlag"]]
  
  c(DescStat(d[b, "RelFL"]), FLFlag = sum(!b), recursive = TRUE)
  
}, FUN.VALUE = numeric(8L)))


catch$LenStatsTagged <- t(vapply(catch[["Data"]], FUN = function(d) {
  
  b <- !d[["RelFLFlag"]] & (is.na(d[["TagAction"]]) | d[["TagAction"]] %in% '1')
  
  c(DescStat(d[b, "RelFL"]), FLFlag = sum(!b), recursive = TRUE)
  
}, FUN.VALUE = numeric(8L)))

catch$LenStatsSex <- lapply(catch[["Data"]], FUN = function(d) {
  
  # b <- d[["RelFLFlag"]]
  
  d$RelSex <- lkp_striper$sex(d[["RelSex"]], asFactor = TRUE)
  
  dd <- Split(
    data = d,
    subset = !RelFLFlag,
    vars = RelFL,
    splitVars = RelSex
  )
  
  t(vapply(dd[["Data"]], FUN = function(x) {
    c(DescStat(x[["RelFL"]]), recursive = TRUE)
  }, FUN.VALUE = numeric(7L)))[c('M', 'F'), ]
  
})

# ******************************************************************************
# might need to limit lower value to compare accross years since measuring
# criteria has changed (and we began tagging sublegals starting ~2010)

# t(vapply(catch[["Data"]], FUN = function(d) {
#   
#   b <- d[["RelFL"]] >= 42
#   
#   c(DescStat(d[b, "RelFL"]), LT42 = sum(!b), recursive = TRUE)
#   
# }, FUN.VALUE = numeric(8L)))
# ******************************************************************************

```

```{r age-stats, warning=FALSE}

catch$AgeStats <- t(vapply(catch[["Data"]], FUN = function(d) {
  
  b <- d[["RelAge"]] >= 2
  
  c(DescStat(d[b, "RelAge"]), recursive = TRUE)
  
}, FUN.VALUE = numeric(7L)))


catch$AgeStatsSex <- lapply(catch[["Data"]], FUN = function(d) {
  
  # b <- d[["RelFLFlag"]]
  
  d$RelSex <- lkp_striper$sex(d[["RelSex"]], asFactor = TRUE)
  
  dd <- Split(
    data = d,
    subset = RelAge >= 2,
    vars = RelAge,
    splitVars = RelSex
  )
  
  t(vapply(dd[["Data"]], FUN = function(x) {
    c(DescStat(x[["RelAge"]]), recursive = TRUE)
  }, FUN.VALUE = numeric(7L)))[c('M', 'F'), ]
  
})

```

<!-- length & age frequency (overall & by sex) -->
<!-- length at age (VBGM) -->

## Effort

Effort data prior to mid-1990 are not available electronically. As time allows, we are entering these data into our current database. Below we summarized currently electronically available data.

#### Days Fished

Annual number of field days for fyke traps (`FT`) hovered around 20 (figure below, top panel). We fished gill nets (`GN`) only one day in the 2010s.

```{r plot-sb-days-fished}

par(mar = c(0.25, 1.9, 0.25, 0.9), oma = c(2, 1, 0.5, 0.45))
mat <- matrix(data = c(1:2), nrow = 2, ncol = 1, byrow = TRUE)
lo <- layout(mat = mat, widths = 1, heights = c(1, 1))
# layout.show(n = lo)

days_sampled <- c("DaysFT", "DaysGN")

# range(effort[["Hours"]][, days_sampled])

p_sb_days <- lapply(days_sampled, FUN = function(v) {
  
  p <- Plot(
    x = effort[["RelYear"]],
    y = effort[["Hours"]][, v],
    adjUsr = 0.5
  )
  
  points(
    p,
    col = grey(0.05, alpha = 0.5),
    type = "h",
    lend = 1,
    lwd = 15,
    grid = TRUE,
    xRng = TRUE
  )
  
  # add y-axis tick labels
  Axis(p, side = 2, labelAdj = 0.4)
  
  lbl <- strsplit(v, split = "Days")[[1]][2]
  xx <- p[["xrng"]][[2]] * 1.0005
  
  # need text() to use srt parameter, not allowed in mtext()
  text(x = xx, y = 20, labels = lbl, srt = -90, xpd = TRUE)
  
  p
})

Axis(p_sb_days[[1]], side = 1, labelAdj = 0.3, interval = 5, offset = 1)
mtext(text = "Year", side = 1, line = 1.20)

layout(mat = 1)
mtext(text = "Days sampled", side = 2, line = 1.5)

# section clean up
rm(p_sb_days, lo, mat, days_sampled)

```

#### Hours Fished

<!-- what is point being plotted; explain data -->

We calculated average hours fished (&pm;SD) given available effort data (figure below). Since roughly 2000, concerns over ESA catch limited fishing fyke traps (`FT`) to no more than 24 hours between checks. On average, gill nets (`GN`) fished between 30 and 55 minutes per set (0.5 to 0.9 hours).

```{r plot-sb-hours-fished}

par(mar = c(0.25, 1.9, 0.25, 0.9), oma = c(2, 1, 0.5, 0.45))
mat <- matrix(data = c(1:2), nrow = 2, ncol = 1, byrow = TRUE)
lo <- layout(mat = mat, widths = 1, heights = c(1, 1))
# layout.show(n = lo)

daily_avg <- list(
  c(Avg = "HoursAvgFT", Var = "HoursVarFT", nm = "FT"),
  c(Avg = "HoursAvgGN", Var = "HoursVarGN", nm = "GN")
)

p_daily_hours <- lapply(daily_avg, FUN = function(v) {
  
  p <- Plot(
    x = effort[["RelYear"]],
    y = effort[["Hours"]][, v[["Avg"]]],
    yerr = sqrt(effort[["Hours"]][, v[["Var"]]]),
    adjUsr = 0.5
  )
  
  lines(p, col = "grey30", lwd = 1.75, lend = 1, xRng = TRUE)
  
  points(
    p,
    type = "p",
    pch = 21,
    col = "white",
    # lwd = 1,
    bg = "grey30",
    cex = 1,
    grid = FALSE
  )
  
  # add y-axis tick labels
  Axis(p, side = 2, labelAdj = 0.4)
  
  # lbl <- strsplit(v, split = "Days")[[1]][2]
  xx <- p[["xrng"]][[2]] * 1.0005
  yy <- median(p[["yrng"]])
  
  # need text() to use srt parameter, not allowed in mtext()
  text(x = xx, y = yy, labels = v[["nm"]], srt = -90, xpd = TRUE)
  
  p
})

Axis(p_daily_hours[[1]], side = 1, labelAdj = 0.3, interval = 5, offset = 1)
mtext(text = "Year", side = 1, line = 1.20)

layout(mat = 1)
mtext(text = "Hours", side = 2, line = 1.5)

# section clean up
rm(p_daily_hours, lo, mat, daily_avg)

```

## Catch

```{r sb-tagged}

# for ease of getting proportion tagged by capture method
tagged <- matrix(
  catch[["Count"]][, "Tagged"],
  ncol = 2,
  dimnames = list(NULL, c("FT", "GN"))
)

```

```{r test-code-delete, eval=FALSE}

# p_catch <- Plot(
#   x = catch[["RelYear"]],
#   y = catch[["Count"]],
#   subset = catch[["CapMethod"]] %in% "GN"
# )
# 
# p_catch$grid(xRng = TRUE)
# 
# points(
#   x = p_catch$data()[["x"]],
#   y = catch[["Count"]][eval(p_catch$subetted), "Total"],
#   type = "h", lwd = 5, lend = 1
# )
# 
# points(
#   x = p_catch$data()[["x"]],
#   y = catch[["Count"]][eval(p_catch$subetted), "Tagged"],
#   type = "p", col = 2, lend = 1
# )

```

```{r plot-tagged, eval=FALSE, fig.asp=1, fig.width=7}

# TODO: plot needs some massaging (05-Feb-2020)
# TODO: see alternate plot below - liking that one better

par(mar = c(0.25, 1.9, 0.25, 0.9), oma = c(2, 1, 0.5, 0.45))
mat <- matrix(data = c(1:3), nrow = 3, ncol = 1, byrow = TRUE)
lo <- layout(mat = mat, widths = 1, heights = c(0.5, 0.8, 0.8))
# layout.show(n = lo)

p_ratio_tagged <- Plot(
  x = range(catch[["RelYear"]]),
  y = c(0, 1),
  adjUsr = 0.5
)

# for grid & denoting y-axis midpoint
p_ratio_tagged$grid(xRng = TRUE)
abline(h = 0.5, col = "steelblue")

# display ratio of tagged via gill net (GN)
points(
  x = unique(catch[["RelYear"]]),
  y = prop.table(tagged, margin = 1)[, "GN"],
  col = "orange3"
)

Axis(p_ratio_tagged, side = 2, labelAdj = 0.3)

plot_num_tagged <- lapply(c("FT", "GN"), FUN = function(v) {
  
  b <- catch[["CapMethod"]] %in% v
  
  p <- Plot(
    # x = catch[b, "RelYear"],
    # y = catch[b, "Count"],
    x = range(catch[["RelYear"]]),
    y = range(catch[["Count"]]),
    adjUsr = 0.5
  )
  
  p$grid(xRng = TRUE)
  
  points(
    # x = p$data()[["x"]],
    x = catch[b, "RelYear"],
    y = catch[["Count"]][b, "Tagged"],
    type = "h",
    col = grey(0.05, alpha = 0.5),
    lend = 1,
    lwd = 8,
  )
  
  # add y-axis tick labels
  pp <- Axis(p, side = 2, labelAdj = 0.4, format = TRUE)
  
  xx <- p[["xrng"]][[2]] * 1.001
  yy <- median(p[["yrng"]])
  
  # need text() to use srt parameter, not allowed in mtext()
  text(x = xx, y = yy, labels = v, srt = -90, xpd = TRUE)
  
  p$AT <- pp$AxisTitle
  p
})

# to show year on define x-axis ticks
Axis(
  plot_num_tagged[[1]],
  side = 1,
  labelAdj = 0.3,
  interval = 5,
  offset = 1
)

# to show x-axis label
mtext(text = "Year", side = 1, line = 1.20)

# to reset layout
layout(mat = 1)

# to show y-axis lable
mtext(
  text = plot_num_tagged[[1]]$AT(var = "Total tagged"),
  side = 2,
  line = 1.3
)

# section clean up
rm(lo, mat, plot_num_tagged, p_ratio_tagged, tagged)

```

Up unitl 2000 and for years when both gear types were fished, gill net (`GN`) contribution to total tagged was between 60% and 80% (figure below, top panel). As a general trend, the annual number of Striped Bass tagged declined.

```{r plot-tagged-alt}

# a possibility
# plot(unique(catch$RelYear), rowSums(tagged), type = "h")

# for adjusting colors using rgb()
# https://www.rapidtables.com/web/color/RGB_Color.html
# c(255, 140, 0) / 256
# c(70, 130, 180) / 256

# for bar colors in plotting num tagged by capture method
lkp_col <- c(
  FT = rgb(red = 0.27, green = 0.50, blue = 0.70, alpha = 0.5),
  GN = rgb(red = 0.99, green = 0.54, blue = 0, alpha = 0.4)
)

par(mar = c(0.25, 1.9, 1.0, 0.4), oma = c(2, 1, 0.5, 0.1))
mat <- matrix(data = c(1:2), nrow = 2, ncol = 1, byrow = TRUE)
lo <- layout(mat = mat, heights = c(0.65, 1))
# layout.show(n = lo)

p_ratio_tagged <- Plot(
  x = range(catch[["RelYear"]]),
  y = c(0, 1),
  adjUsr = 0.5
)

# for grid & denoting y-axis midpoint
p_ratio_tagged$grid(xRng = TRUE)
# abline(h = 0.5, col = "steelblue")
abline(h = 0.5, col = grey(level = 0.5, alpha = 0.5), lwd = 3)

# display ratio of tagged via gill net (GN)
points(
  x = unique(catch[["RelYear"]]),
  y = prop.table(tagged, margin = 1)[, "GN"],
  cex = 0.75,
  pch = 1,
  col = "grey30"
)

Axis(
  p_ratio_tagged,
  side = 2,
  labelAdj = 0.3
)

mtext(text = "GN fraction", side = 2, line = 1.6)

# to minimize gap between plots
par(mar = c(0.25, 1.9, 0.25, 0.4))

# for empty plot based on annual number tagged
p_tagged <- Plot(
  x = catch[["RelYear"]],
  y = catch[["Count"]][, "Tagged"],
  adjUsr = 0.5
)

# add bars representing number tagged, shaded by capture method
points(
  p_tagged,
  type = "h",
  col = lkp_col[catch[["CapMethod"]]],
  lwd = 8,
  xRng = TRUE
)

# to show year on define x-axis ticks
Axis(
  p_tagged,
  side = 1,
  labelAdj = 0.3,
  interval = 5,
  offset = 1
)

# for formatting of yaxis 10^3
yaxis_tagged <- Axis(
  p_tagged,
  side = 2,
  labelAdj = 0.3,
  format = TRUE
)

# to show x-axis label
mtext(text = "Year", side = 1, line = 1.20)

# to show y-axis lable
mtext(
  text = yaxis_tagged$AxisTitle(var = "Total tagged"),
  side = 2,
  line = 1.3
)

# to reset plot window for legend placement
layout(mat = 1)

legend(
  x = p_tagged[["xrng"]][2],
  y = p_tagged[["yrng"]][2] * 1.07,
  legend = names(lkp_col),
  fill = lkp_col,
  border = NA,
  bty = "n",
  ncol = 2,
  xjust = 1,
  xpd = TRUE
)

# section clean up
rm(lo, mat, yaxis_tagged, lkp_col)

```

#### Sex Ratio

We assigned sex based on the presence (&male;) or absence (&female;) of milt. We calculated sex ratio as a fration of total caught (including unassigned; figure below). This is by no means a comprehensive look at sex ratio and more in-depth analyses are required, but it does demonstrate some shift over time. *Note*: n=17 Striped Bass for 2011 Gill net operations (one day of sampling).

```{r plot-sex-ratio}

lkp_col_sex_ratio <- c(
  FT = rgb(red = 0.27, green = 0.50, blue = 0.70, alpha = 0.9),
  GN = rgb(red = 0.99, green = 0.54, blue = 0, alpha = 0.9)
)

par(mar = c(0.5, 2.0, 1, 0.45), oma = c(2, 1, 0.5, 0.2))

p_sex_ratio <- Plot(
  x = catch[["RelYear"]],
  y = catch[["SexRatio"]][, c("PropM", "PropF")], adjUsr = 0.5
)

p_sex_ratio$grid(xRng = T)

points(
  x = catch[["RelYear"]],
  y = catch[["SexRatio"]][, "PropM"],
  col = lkp_col_sex_ratio[catch[["CapMethod"]]],
  pch = "\u2642"
)

points(
  x = catch[["RelYear"]],
  y = catch[["SexRatio"]][, "PropF"],
  col = lkp_col_sex_ratio[catch[["CapMethod"]]],
  pch = "\u2640"
)

Axis(p_sex_ratio, side = 1, labelAdj = 0.2, interval = 5, offset = 1)
Axis(p_sex_ratio, side = 2, labelAdj = 0.4)
mtext(text = "Year", side = 1, line = 1.2)
mtext(text = "Ratio", side = 2, line = 1.9)

legend(
  x = p_sex_ratio[["xrng"]][2],
  y = p_sex_ratio[["yrng"]][2] * 1.10,
  legend = names(lkp_col_sex_ratio),
  fill = lkp_col_sex_ratio,
  border = NA,
  bty = "n",
  ncol = 2,
  xjust = 1,
  xpd = TRUE
)

```

## Tag Returns

A major component of this mark-recapture study is angler tag returns. Every angler who returns a disc tag receives a commendation card and reward (separately and if applicable). A commendation card thanks the angler for contributing to the "betterment of fishing in California" and provides a bit of information about the fish (e.g., date and location tagged).

We tend to process angler tag returns in batches (i.e., several to half dozen at a time). Rewards are paid within six to eight weeks of processing. Total annual payments vary but for the last 10 years mostly have been between \$1,000 and \$2,000 (table below). Year represents period in which the tag was processed, not return or release year.

*Note1*: We only have tag processing data-time from ~October 2011 to present. Tags prior to which were processed but have no associated date-time stamp.

*Note2*: Data for 2011 may not include all tags processed that year.

*Note3*: We are still processing tags returned in 2020.

<!-- address $10 tag -->
<!-- address why so few in 2016 -->

```{r returns-relyear}

returns <- Split(
  data = Tagging[["TagReturns"]],
  subset = ,
  vars = c(
    RelDate, RelLoc, CapMethod, TagNum,
    RelVal, RetDate, FlaggedRetDate, RetYear
  ),
  splitVars = RelYear
)

# for desired data type
returns$RelYear <- as.numeric(returns[["RelYear"]])

```

```{r returns-var}

returns$Count <- vapply(
  returns[["Data"]],
  FUN = nrow,
  FUN.VALUE = numeric(1L),
  USE.NAMES = FALSE
)

returns$FirstYear <- vapply(returns[["Data"]], FUN = function(d) {
  sum(d[["RetYear"]] %in% 1)
}, FUN.VALUE = numeric(1L), USE.NAMES = FALSE)

returns$Payment <- vapply(returns[["Data"]], FUN = function(d) {
  b <- grepl(pattern = "\\d", x = d[["RelVal"]])
  sum(as.numeric(d[b, "RelVal"]))
}, FUN.VALUE = numeric(1L), USE.NAMES = FALSE)

returns$PaymentFY <- vapply(returns[["Data"]], FUN = function(d) {
  b <- grepl(pattern = "\\d", x = d[["RelVal"]]) &
    d[["RetYear"]] %in% 1
  sum(as.numeric(d[b, "RelVal"]))
}, FUN.VALUE = numeric(1L), USE.NAMES = FALSE)

returns$ByRelVal <- t(vapply(returns[["Data"]], FUN = function(d) {
  lvls <- names(lkp_striper[["tagvalue"]])
  table(factor(d[["RelVal"]], levels = lvls))
}, FUN.VALUE = numeric(8L)))

returns$FirstYearNR <- vapply(returns[["Data"]], FUN = function(d) {
  sum(d[["RetYear"]] %in% 1 & grepl(pattern = "\\D", x = d[["RelVal"]]))
}, FUN.VALUE = numeric(1L), USE.NAMES = FALSE)

```

```{r testing, eval=FALSE}

# a possibility for output (maybe heatmap)

lapply(returns[["Data"]], FUN = function(d) {
  y <- format(d[["RetDate"]], format = "%Y")
  table(as.numeric(y), useNA = "always")
})

```

```{r reward-payment}

reward <- factor(
  Tagging[["RewardPayment"]][["Reward"]],
  levels = c("10", "20", "50", "100", "Payment"),
  labels = c(paste0("$", c("10", "20", "50", "100")), "Payment")
)

tbx <- xtabs(
  formula = NumProc ~ ProcYear + reward,
  data = Tagging[["RewardPayment"]]
)

dol <- aggregate(
  formula = Payment ~ ProcYear,
  data = Tagging[["RewardPayment"]],
  FUN = sum
)

tbx[, "Payment"] <- paste("$", (format(dol[["Payment"]], big.mark = ",")))

knitr::kable(tbx, format = "markdown", align = "r")

# section clean up
rm(dol, tbx, reward)

```

We classify tag returns based on a tag's time at large. That is, a tag returned within a year of release (i.e., tagging) is a first-year return, a tag returned between one and two years of release is a second-year return, and so on. Primarly, we use first-year and second-year returns for some metrics to mitigate against bias (e.g., angler didn't report or recall capture date; angler returned the tag many years after capture).

We calculate return rate simply by dividing number returned by number released. Below, we display first-year non-reward return rate. Here we do not correct for non-response --- a process that uses reward return rates --- but the trend is markedly downward over the study's history.

```{r plot-return-rate}

# ******************************************************************************
# both are equivalent but subset does require parens around `TagAction`; both
# get the number of non-reward tags released per year

# aggregate(
#   formula = Count ~ RelYear,
#   # formula = Count ~ RelYear + TagValue,
#   data = Tagging$StripedBass,
#   FUN = sum,
#   subset = (TagAction %in% 1 | is.na(TagAction)) &
#     TagValue %in% c("NM", "NR")
# )[["Count"]]
# 
# unname(rowSums(tag_values[, c("NM", "NR")]))
# ******************************************************************************

# for annual count of non-reward tags released
non_reward_tags_rel <- unname(rowSums(tag_values[, c("NM", "NR")]))

par(mar = c(0.5, 2.3, 0.25, 0.45), oma = c(2, 1, 0.5, 0.45))

# annual return rate
p_return_rate <- Plot(
  x = returns[["RelYear"]],
  y = returns[["FirstYearNR"]] / non_reward_tags_rel,
  adjUsr = 0.5
)

points(p_return_rate, xRng = TRUE)
# points(p_return_rate, xRng = FALSE)
Axis(p_return_rate, side = 1, labelAdj = 0.3, interval = 5, offset = 1)
# Axis(p_return_rate, side = 1, labelAdj = 0.3, offset = 1)
Axis(p_return_rate, side = 2, labelAdj = 0.4)
mtext(text = "Year", side = 1, line = 1.2)
mtext(text = "Return rate", side = 2, line = 2.2)

# section clean up
rm(non_reward_tags_rel)

```

```{r plot-tagged-sub, eval=FALSE}

# plots fork length of fish tagged color coded to denote legal & sub-legal

par(mar = c(2.5, 2.3, 1.2, 0.45), oma = c(1.1, 1, 0.5, 0.45))

lkp_col <- c(
  `TRUE` = rgb(red = 0, green = 0.35, blue = 0.65, alpha = 0.1),
  `FALSE` = rgb(red = 0, green = 0, blue = 0, alpha = 0.1)
)

p_tagged_sub <- Plot(
  Tagging[["StripedBass"]],
  x = RelYear,
  y = RelFL,
  subset = RelYear > 2009 & CapMethod %in% 2 & !RelFLFlag &
    # RelFL < 42 &
    (is.na(TagAction) | TagAction %in% '1')
)

p_tagged_sub$grid(xRng = TRUE)
Axis(p_tagged_sub, side = 2, labelAdj = 0.4)
Axis(p_tagged_sub, side = 1, labelAdj = 0.3)

# # points(p_tagged_sub)
points(
  x = jitter(p_tagged_sub$data()[["x"]]),
  y = p_tagged_sub$data()[["y"]],
  pch = 20,
  cex = 0.75,
  # col = rgb(red = 0, green = 0, blue = 0, alpha = 0.1)
  col = lkp_col[as.character(p_tagged_sub$data()[["y"]] < 42)]
)

# sub() employeed to change alpha so point is visible
legend(
  x = p_tagged_sub[["xrng"]][[2]], 
  y = p_tagged_sub[["yrng"]][[2]] * 1.11,
  legend = c("sub-legal", "legal"),
  col = sub(pattern = "1A$", replacement = "CC", x = lkp_col),
  pt.cex = 2,
  pch = 20,
  border = NA,
  bty = "n",
  ncol = 2,
  xpd = TRUE,
  xjust = 0.75
)

mtext(text = "Year (fraction sub-legal)", side = 1, line = 2.2)
mtext(text = "Fork length (cm)", side = 2, line = 2)

# **********************************************************

# to calculate fraction tagged of sub-legal sized fish
frac <- aggregate(
  formula = y ~ x,
  data = p_tagged_sub$data(),
  FUN = function(l) mean(l < 42)
)

# to display fraction (`frac`) below year xaxis tick labels
text(
  x = frac[["x"]],
  # y = p_tagged_sub[["yrng"]][[2]] * 1.055,
  y = p_tagged_sub[["yrng"]][[1]] * 0.58,
  # labels = sprintf(fmt = "%.2f", frac[["y"]]),
  labels = sprintf(fmt = "(%.2f)", frac[["y"]]),
  cex = 0.75,
  col = "grey50",
  xpd = TRUE
)

# section clean up
rm(lkp_col, frac)

```

```{r eval=FALSE}


p_loc <- Plot(
  x = range(catch[["RelYear"]]),
  y = range(as.numeric(dimnames(catch[["DaysLoc"]])[[2]])),
  adjUsr = 0.5
)

# p_loc$grid(yRng = F, ny = 1)

Axis(p_loc, side = 1, labelAdj = 0.3)
# Axis(p_loc, side = 2, labelAdj = 0.3)
axis(side = 2, at = 1:16, labels = 1:16)

#

b <- catch[["CapMethod"]] %in% "GN"

dn <- dimnames(catch[b, "DaysLoc"])

# plot(c(1969, 2019), c(1, 16), type = "n")
# rect(xleft = 1968.5, ybottom = 0.75, xright = 1969.5, ytop = 1.25, border = NA, col = rgb(red = 0, green = 0.35, blue = 0.65, alpha = 0.5))

# # catch[[b, "DaysLoc"]]
# # 
# rect(
#   xleft = catch[b, "RelYear"] - 0.5,
#   ybottom = (1:16) - 0.25,
#   xright = catch[b, "RelYear"] + 0.5,
#   ytop = (1:16) + 0.25,
#   border = NA,
#   col = rgb(red = 0, green = 0.35, blue = 0.65, alpha = 0.5)
# )

# catch[b, "RelYear"] - 0.5
# 
# # 
# 
test <- lapply(dn[[1]], FUN = function(x, l) {

  y <- gsub(pattern = "\\D", replacement = "", x = x)
  y <- as.numeric(y)
  
  x1 <- y - 0.5
  x2 <- y + 0.5
  
  lapply(l, FUN = function(ll) {
    
    yy <- as.numeric(ll)
    
    a <- catch[["DaysLoc"]][x, ll] / 52
    
    # col <- "grey85"
    col <- "white"
    
    if (a > 0) 
      col <- rgb(red = 0, green = 0.35, blue = 0.65, alpha = a)
    
    rect(
      xleft = x1,
      ybottom = yy - 0.5,
      xright = x2,
      ytop = yy + 0.5,
      border = NA,
      col = col
    )
    
  })


}, l = dn[[2]])
p_loc$yticks()
p_loc$ticks(side = 2)

```

```{r plot-days-loc, fig.asp=0.5, fig.width=8}

# chunk creates 'heatmap' for days sampled by location per gear; creates layout
# to make possible the display of a color bar for both gear types

# note: alpha in `rgb()` uses sqrt(days / 100), this works because no value is >
# 70 but it will fail say if days > 100

# TODO: it is quite possible Map() & lapply() are not needed here, just
# experiment with vectorization using rect() or segements() to plot
# catch[["DaysLoc"]] matrix; will keep the way it is for now (21-Feb-2020) to
# move on to other items in this review

# create layout for main plot & color bar (another plot)
lo <- layout(
  mat = matrix(data = 1:2, ncol = 2, nrow = 1, byrow = TRUE),
  widths = c(5.5, 0.25)
)

# layout.show(lo)

# establish inner & outer margins for main plot
par(mar = c(2, 2, 0.25, 2), oma = c(0.5, 0.5, 0.25, 1))

# main plot based on release year range & numeric location range
p_loc <- Plot(
  x = range(catch[["RelYear"]]),
  y = range(as.numeric(dimnames(catch[["DaysLoc"]])[[2]])),
  adjUsr = 0.5
)

# for use in axis below
locs <- seq_len(p_loc[["yrng"]][[2]])

# establish grid, axes ticks, and axes labels ************************
p_loc$grid(xRng = TRUE, yRng = TRUE)

Axis(p_loc, side = 1, labelAdj = 0.3, interval = 5, offset = 1)

axis(
  side = 2,
  at = locs,
  labels = sub(pattern = " ", replacement = "0", x = format(locs)),
  las = 1,
  tcl = -0.2,
  col = "transparent",
  col.ticks = "grey50",
  col.axis = "grey50"
)

mtext(text = "Year", side = 1, line = 1.2)
mtext(text = "Location code", side = 2, line = 1.3)

# to plot value (i.e., days at location) as rectangle for heatmap appearance
p_loc_heat <- Map(f = function(y, g) {
  
  # establish left & right sides of rectangle
  x1 <- y - 0.5
  x2 <- y + 0.5
  
  # for each location add a rectangle accordingly
  lapply(locs, FUN = function(l) {
    
    # to get correct value from catch[["DaysLoc"]] matrix
    x <- paste(y, g, sep = ".")
    
    # attemps to programatically generate value for alpha in rgb()
    # a <- catch[["DaysLoc"]][x, as.character(l)] / 69
    # a <- (log10(catch[["DaysLoc"]][x, as.character(l)] + 1)) %% 1
    a <- sqrt(catch[["DaysLoc"]][x, as.character(l)] / 100)
    
    # default color 
    col <- NA
    
    # color when greater than 0 & depending on gear type
    if (a > 0)
      if (g %in% "FT")
        col <- rgb(red = 0.27, green = 0.50, blue = 0.70, alpha = a)
      if (g %in% "GN")
        col <- rgb(red = 0.99, green = 0.54, blue = 0, alpha = a)
    
    # display the rectangle with appropriate color or no color (NA)
    rect(
      xleft = x1,
      ybottom = l - 0.4,
      xright = x2,
      ytop = l + 0.4,
      border = NA,
      col = col
    )
    
    # output color for use in color bar
    col
  })
  # end inner lapply()

}, y = catch[["RelYear"]], g = catch[["CapMethod"]])

# establish second plot - which is the color bar
par(mar = c(2.5, 0, 2.5, 0.1))

# 1 & 2 for gear types, 70 used as max because max days at one loc = 69
p_loc_color_bar <- Plot(x = c(1, 2), y = c(1, 70))

Axis(p_loc_color_bar, side = 2, labelAdj = 0.3, cexAxis = 0.8)

# for labeling color bar with gear type - easier this way than with axis()
mtext(
  text = c("FT", "GN"),
  side = 3,
  at = c(1.1, 1.9),
  line = -0.5,
  cex = 0.8,
  col = "grey50"
)

# y-label for color bar
mtext(text = "Days sampled", side = 2, line = 1.1, cex = 0.8)

# for use in segments()
y0 <- seq(from = 0, to = 60, by = 10)
y1 <- seq(from = 10, to = 70, by = 10)
a <- sqrt(y1 / 100)

# y0 <- seq(from = 0, to = 69, by = 1)
# y1 <- seq(from = 1, to = 70, by = 1)
# a <- sqrt(y1 / 100)

# add segements based on increasing color
segments(
  x0 = 1,
  y0 = y0,
  y1 = y1,
  col = rgb(red = 0.27, green = 0.50, blue = 0.70, alpha = a),
  lwd = 25,
  lend= 1
)

segments(
  x0 = 2,
  y0 = y0,
  y1 = y1,
  col = rgb(red = 0.99, green = 0.54, blue = 0, alpha = a),
  lwd = 25,
  lend= 1
)

# section clean up
rm(lo, locs, p_loc_heat, p_loc_color_bar, y0, y1, a)

```

```{r eval=FALSE}

# attempt at color bar for days-at-location plot above
# Thu Feb 20 08:56:35 2020 ------------------------------

plot(factor(c(1, 2)), c(1, 70), type = "n", lty = 0)




Plot(x = c(1, 1), y = c(1, 70))$xticks()

segments(x0 = 0.8, y0 = 0, y1 = 10, col = 1, lwd = 20, lend= 1)
segments(x0 = 0.8, y0 = 10, y1 = 20, col = 2, lwd = 20, lend= 1)

```

